% !TeX root = ../main.tex
\chapter{Common Knowledge}
\chapterauthor{Aden Chen}

If differences in beliefs originate from asymmetric information, then knowing that others hold different beliefs should make a rational agent revise her own belief.
To what extent would this revision happen?
\Textcite{aumann1976AgreeingDisagree} shows that if agents with different information were to ``agree'' on their beliefs --- if their beliefs are common knowledge ---then their beliefs must be exactly the same; agents cannot agree to disagree.

% In this seminal 4-page paper, Aumann writes:
%
% \begin{displayquote}
%   We publish this observation with some diffidence, since once one has the appropriate framework, it is mathematically trivial.
%   Intuitively, though, it is not quite obvious ...
% \end{displayquote}
%
% ``Mathematically trivial'' is surely an understatement, but there is an element of truth: the more important contribution of this paper is perhaps the model of information or knowledge used, and the definition of common knowledge.
% Once this framework is in place, the main result can be derived using a rather simple, geometric argument.
% ``Not quite obvious'' is surely an understatement.

% In what follows, we will walk over the model of information Aumann worked with and his definition of common knowledge.
% With the right definitions, Aumann's result can be recovered with a simple, geometric argument.
% \mnote{
% In this seminal 4-page paper, Aumann writes:
% ``We publish this observation with some diffidence, since once one has the appropriate framework, it is mathematically trivial.
% Intuitively, though, it is not quite obvious...''
% }

To describe the epistemic model Aumann worked with, we start with the finite state space \(\Omega\).
Each element of this set (each state) \(\omega \in \Omega\) is interpreted as a potential world the agents could be living in.
It gives a complete description of the world, including the information each individual possesses and their beliefs.

For players \(i = 1, 2\), let \(\cP_i\) be a partition of \(\Omega\) that reflects how finely each player can distinguish the different potential states of the world using their information.
For a state \(\omega \in \Omega\), write \(\cP_i(\omega)\) for the cell in \(\cP_i\) containing \(\omega\).
If the true state is \(\omega\), player \(i\) would know that the state must be in \(\cP_i(\omega)\), but cannot pin down the exact state if \(\cP_i(\omega)\) contains multiple elements.
% \mnote{
%   Compared to a model of type spaces, this model makes it clearer what exactly each player \emph{knows}.
%   % if we understand \(i\) knowing proposition \(q\) as meaning ``\(i\) believes the true statement \(q\) with certainty.''
% }

Since each \(\omega\) offers a complete description of the way information is distributed to each individual, we can assume that each agent \(i\) knows the partitions of both players \(\cP_i\) and \(\cP_j\), though of course she can observe only \(\cP_i(\omega)\) and not \(\cP_j(\omega)\) at \(\omega\).

\begin{definition}
  Say \(i\) \vocab{knows} event \(E \subset \Omega\) at \(\omega\) if \(E \supset \cP_i(\omega)\).
  Say \(E \subset \Omega\) is \vocab{common knowledge} at \(\omega\) if \(i\) knows \(E\), \(j\) knows \(E\), \(i\) knows \(j\) knows \(E\), \(j\) knows \(i\) knows \(E\), \(i\) knows \(j\) knows \(i\) knows \(E\), ad infinitum.
\end{definition}

To make sense of this definition, consider a proposition \(q\) which can be either true or false, and let \(E\) denote the set of states in which \(q\) is true.
When does agent \(i\) \emph{know} --- believe with certainty --- that \(q\) is true?
She does so exactly when \(q\) is true in each of the states deemed possible by her, i.e., when \(\cP_i(\omega) \subset E\).
\mnote[-10em]{This definition is reminiscent of the idea that knowledge is justified true belief. Justified in the sense of concluding from \(\omega \in \cP_i(\omega) \subset E\), true in the sense that \(E \ni \omega\). Gettier cases are not a problem here since in this model we leave no room for epistemic luck. All logical deductions of the agents are always accurate.}


Write \(\cP_1 \land \cP_2\) for the finest common coarsening of \(\cP_1\) and \(\cP_2\).
We have the following characterization of common knowledge.
\mnote{
  The set of all partitions with the refinement order (\(\cP_1 \leq \cP_2\) if \(\cP_2\) refines \(\cP_1\)) is a lattice.
  \(\cP_1 \land \cP_2\) is the infimum (or the meet) of \(\cP_1\) and \(\cP_2\).
}

\begin{proposition}
  \(E\) is common knowledge at \(\omega\) iff it contains \((\cP_1 \land \cP_2)(\omega)\).
\end{proposition}
\begin{proof}
  At \(\omega\), \(1\) knows that the true state must be in \(\cP_1(\omega)\).
  She can thus infer that \(2\) must have observed \(\cP_2(\omega')\) for some \(\omega' \in \cP_1(\omega)\).
  Thus, \(1\) knows \(2\) knows \(E\) if for each of these \(\omega'\), \(E \supset \cP_2(\omega')\).
  In other words, \(1\) knows \(2\) know \(E\) if \(E\) includes all \(P \in \cP_2\) intersecting \(\cP_1(\omega)\).

  To generalize this idea to higher order beliefs, say \(\omega'\) is \vocab{reachable} \(\omega\) if there is a sequence \(P^1, \dots, P^k\) such that \(\omega \in P^1\), \(\omega' \in P^k\), and consecutive \(P^j\) intersect and belong alternatively to \(\cP_1\) and \(\cP_2\).
  The analysis above shows that the \(k\)th statement ``\(1\) knows \(2\) knows ... \(E\)'' holds iff \(E\) contains all \(\omega'\) reachable from \(\omega\) in step \(k\).
  Thus \(E\) is common knowledge iff it contains the union of each reachable \(\omega'\), for steps \(k \geq 0\).
  This union is precisely \((\cP_1 \land \cP_2)(\omega)\).
\end{proof}

Say event \(E \subset \Omega\) is \vocab{self-evident to \(i\)} if for each \(\omega \in E\) we have \(\cP_i(\omega) \subset E\).
Say \(E\) is \vocab{self-evident} if it is self-evident to each agent \(i\).
It is not hard to see that \(E\) is self-evident to \(i\) iff it can be written as the union of cells in \(\cP_i\).
Thus from the previous result we have the following
\begin{corollary}
  \(E\) is common knowledge iff it contains a self-evident event.
\end{corollary}


Now suppose each player starts with the same prior \(\pi\) over \(\Omega\).
At \(\omega\), upon receiving her information given by \(\cP_i\), player \(i\) updates her belief to \(\pi\left(\cdot \mid \cP_i(\omega)\right)\).

\begin{proposition}
  Fix event \(A \subset \Omega\) and numbers \(p_1, p_2\).
  If it is common knowledge at \(\theta\) that \(\pi(A|\cP_1(\omega)) = p_1\) and \(\pi(A|\cP_2(\omega)) = p_2\), then \(p_1 = p_2\).
\end{proposition}
\begin{proof}
  Write \(P = (\cP_1 \land \cP_2)(\omega) = \cup_j P^j\) where \(P^j \in \cP_1\).
  With the characterization of common knowledge above, we know that \(\pi(A \mid \cP^j) = p_1\) for each \(j\).
  Thus by the law of iterated expectation, \(\pi(A \mid P) = p_1\).
  A symmetric argument shows \(\pi(A|P) = p_2\).
\end{proof}


In this sense, it is impossible to ``agree to disagree''!
So, what do we make of this result when people do seem to agree to disagree?
Some potential explanations:

\paragraph{The common prior assumption may not be true.}
Our model assumes that each player start from the same prior, which is then updated after she observes the cell containing the true state \(\cP_i(\omega)\).
Thus while we assume the same prior, we can model differences in beliefs due to the players having different information.

To relax the common assumption, one should first ask what, if not information, causes the differences in priors?
Without a good answer to this question, explaining behavior based on arbitrary priors seem to some extent like a slippery slope --- one akin to explanations based on preferences.

\paragraph{Rationality.}
Agents might not update their beliefs accurately, or might not trust that others would update their beliefs accurately.

\paragraph{Common knowledge is a strong requirement.}
The definition of knowledge here allows for no uncertainty, and common knowledge requires each agent to know all higher order beliefs hold.
Is this really the notion of agreement that people have when they agree to disagree?

The most likely scenario for common knowledge is probably public announcements (``town hall meetings'').
When information is disseminated in public, not only does everyone knows of the information, they also know that others know this information.
(The public announcement is a \emph{self-evident} event in the sense defined above.)
But even in this case, can the do agents really believe with certainty that all others received the information accurately?

In such cases it is perhaps only case the that agents have ``almost common knowledge,'' where instead of \emph{all} statements of the form ``\(i\) knows \(j\) knows ...'' are true, only a (potentially very large) number of these are true.
In terms of the agents' actions, however, common knowledge can be very different from ``almost common knowledge'' (e.g., \textcite{rubinstein1992ElectronicMail}).


